



```{r}
rm(list=ls())
library(msm)
data_cleaned= read.csv("data_cleaned.csv")
#data_cleaned = read.csv("data_cleaned_simplified.csv")
#data_cleaned = read.csv("sample.csv")

```



### Load msm package:

```{r}
library("msm")
data_msm = data_cleaned
```

### Frequency table of pairs of consecutive states

```{r}
library(xtable)
data_temp = data_cleaned
#data_temp$state[data_temp$state == 1] = "HEALTHY"
#data_temp$state[data_temp$state == 2] = "WP"
#data_temp$state[data_temp$state == 3] = "PM/DEAD"

data_frequency_table = statetable.msm(state,coral,data_temp)
data_frequency_table
print(xtable(data_frequency_table,type='latex'))

               
```







```{r}

statetable.msm(state,coral,data_msm)
Q <- rbind ( c(1,1,1 ),c(1,1,1),c(0,0,0))
print("Transition indicator matrix")
print(Q)
# make partial mortality combined with dead
# optional
Q.crude <- crudeinits.msm(state ~months, coral, data=data_msm, qmatrix=Q)
Q.crude
```



```{r}
coral.msm <- msm( state ~ months, subject=coral, data = data_msm,control=list(maxit = 500,fnscale=5000,trace = 1, report = 1),qmatrix = Q.crude)
coral.msm
```



```{r}
#coral.msm <- msm( state ~ months, subject=coral, data = data_msm,  control=list(maxit = 1000,fnscale=500,trace = 1, report = 1,reltol = 1e-8),qmatrix = Q.crude,covariates = ~ bur_coral + bl_coral+control,covinits = list(bur_coral = c(1,1,0,0),bl_coral = c(1,1,0,0)),fixedpars = c(7,8,11,12,17,20,21,24),pci=c(5,10))

#coral.msm <- msm( state ~ months, subject=coral, data = data_msm,  control=list(maxit = 1000,fnscale=500,trace = 1, report = 1,reltol = 1e-8),qmatrix = Q.crude,covariates = ~ bur_coral + bl_coral+control,pci = c(5,10))
#fixed_par = c(7,8,11,12,19,20,23,24,27,28,31,32,35,36,39,40,43,44)


#coral.msm <- msm( state ~ months, subject=coral, data = data_msm,  control=list(maxit = 1000,fnscale=500,trace = 1, report = 1,reltol = 1e-8),qmatrix = Q.crude,covariates = ~ bur_coral + bl_coral+control,covinits = list(bur_coral = c(1,1,0,0),bl_coral = c(1,1,0,0)),fixedpars = fixed_par,pci=c(3,4,5,6,7,8,9))

#coral.msm <- msm( state ~ months, subject=coral, data = data_msm,  control=list(maxit = 1000,fnscale=500,trace = 1, report = 1,reltol = 1e-8),qmatrix = Q.crude,covariates = ~ bur_coral + bl_coral+control,covinits = list(bur_coral = c(1,1,1,1),bl_coral = c(1,1,1,1)),pci = c(10))


coral.msm <- msm( state ~ months, subject=coral, data = data_msm,  control=list(maxit = 1000,fnscale=500,trace = 1, report = 1,reltol = 1e-8),qmatrix = Q.crude,covariates = ~ bur_coral + bl_coral+control,covinits = list(bur_coral = c(1,1,1,1),bl_coral = c(1,1,1,1)),pci = c(5,10,15))

coral.msm
```

```{r}


# baseline hazard for "HEALTHY" to "WP"
res = 0.006406*18.0971
upper = 0.01135*37.5427
lower = 0.003614*8.72352


# baseline hazard for "HEALTHY" to "PM/DEAD"
res = summary(coral.msm)
baseline=c(0.026328,0.016826, 0.04119)
baseline*res$hazard$`timeperiod[5,10)`[2,]
baseline*res$hazard$`timeperiod[10,15)`[2,]
baseline*res$hazard$`timeperiod[15,Inf)`[2,]








```



```{r}
#summary(coral.msm)

res = summary(coral.msm)
res$hazard
a = res$hazard$bl_coral
#a[a==1] = '/'
print(xtable(a,type='latex'))

#write.csv(a,"~/Downloads/test.csv")
b =res$hazard$bur_coral
#b[b==1] = '/'
#write.csv(b,"~/Downloads/test2.csv")
print(xtable(b,type='latex'))
c = res$hazard$control
#c[c==1] = '/'
print(xtable(c,type='latex'))
#write.csv(c,"~/Downloads/test3.csv")

d = res$hazard$`timeperiod[5,10)`
#d[d==1] = '/'
print(xtable(d,type='latex'))
#write.csv(d,"~/Downloads/test4.csv")

e = res$hazard$`timeperiod[10,Inf)`
#e[e==1] = '/'
#write.csv(e,"~/Downloads/test5.csv")
print(xtable(e,type='latex'))

```





```{r}
pmatrix.msm(coral.msm, t=5)
```


#### Mean sojourn times

The function sojourn.msm extracts the estimated mean sojourn times in each transient state r, for a given set of covariate values.

```{r}
sojourn.msm(coral.msm)
```


# Hazard ratio of the bleaching, buried, and control covariates



```{r}
hazard.msm(coral.msm)
```


# Survival plot

an important use of multi-state models is in predicting the probability of survival for patients in increasingly severe states of disease, for some time t in the future. This can be obtained directly from the transition probability matrix P(t).

```{r}
plot(coral.msm, legend.pos=c(8, 1))
```

# Goodness of fit
# Observed and expected prevalence
```{r}
prevalence.msm(coral.msm,times = seq(0,20,1))
plot.prevalence.msm(coral.msm,mintime = 0, maxtime = 20,legend.pos = c(0,100))
```
# Pearson-type goodness-of-fit


# cut the interval and redo.

```{r}
pearson.msm(coral.msm)
```






